Figures and Analysis for 2023 Manuscript

### 2023-12-01 data-raw/Manuscript2023_Workflow.Rmd
Made sure not to evaluate chunks

```{r pooled_fitting_choices}
# Types of Choices:
## Dose Normalized
## Log10 transform
## Scale time

@@

```{r joint_fitting_choices}
# Joint
my_pk_000j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

@@

```{r Evaluation-helpers}
# Evaluation
# Current:

@@

```{r Evaluating choices}

@@

```{r load_multirefs}
# Reference Multiples
multiRef_ChemSpec <- cvt_df %>%
  ungroup() %>%
  distinct(analyte_dtxsid, species, document_id) %>%
  group_by(analyte_dtxsid, species) %>%
  count() %>%
  arrange(desc(n)) %>%
  filter(n > 1) %>%
  dplyr::select(!n)
multiRef_ChemSpec %>% nrow() # Surprisingly only 13 Chem + Species have multiple references

@@

```{r Plotting_Summary}
finalist_options <- c("000 Pooled", "000 Joint",
                      "100 Pooled", "100 Joint",
                      "101 Pooled", "101 Joint")

@@

```{r Final_Options}
my_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species, Reference))

@@

```{r data_aggregation}
# Need the following data for any of the subsequent plots

@@

### Figure 4A: Replicated timepoint variation in CvTdb
```{r fig4A}

@@

```{r fig4A-2-alt}
iad_summary <- cvt %>% filter(conc > 0, n_subjects_in_series == 1) %>%
  group_by(analyte_dtxsid, species, administration_route_normalized,
           conc_medium_normalized, dose_level_corrected, document_id,
           time_hr) %>%
  filter(n() > 1) %>%
  summarize(conc_sd_calc = sd(conc),
            conc_avg_calc = mean(conc),
            n_subjects_in_series = n())

@@

```{r fig4B-cowplot}

@@

```{r cvt_analysis}
data_range <- all_my_data %>%
  filter(Detect) %>%
  group_by(Chemical, Species, Reference, Media, Route, Dose, Dose.Units) %>%
  summarize(maxT = max(Time)/24,
         concRange = log10(max(Conc)/min(Conc))) %>%
  ungroup()

@@

```{r sFig2-model_performance}
my_preds %>%
  right_join(winmodel) %>%
  rowwise() %>%
  mutate(Pred_Obs = Conc_trans/Conc_est) %>%
  filter(is.finite(Pred_Obs),
         model %in% c("model_1comp", "model_2comp")) %>%
  # mutate(pass = ifelse(Pred_Obs < 0.5, TRUE, FALSE)) %>%
  # pull(pass) %>% table()
  ggplot(aes(x = Pred_Obs)) +
  geom_histogram(binwidth = 0.5,
                 fill = "grey5") +
  scale_x_log10(limits = c(NA, 1E6)) +
  expand_limits(y = 0) +
  geom_vline(xintercept = c(0.5, 2), color = "red3", linetype = "dashed") +
  facet_grid(cols = vars(model),
             rows = vars(Route)) +
  labs(x = "Dose-normalized\nPredicted/Observed",
       y = "Number of Observations") +
  theme_classic() +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.4))

@@

```{r fig5-modelPerform-v-dataVar}

@@

```{r fig5Alt}
# Redo combined summary
combined_summary <- bind_rows(sud_summary, iad_summary) %>%
  rowwise() %>%
  mutate(twofold_95 = ifelse((conc_avg_calc + (2*conc_sd_calc))/conc_avg_calc < 2,
                             TRUE, FALSE))

@@

```{r fig6-goodness-of-fit}

@@

```{r sFig3-sigmaDistribution}

@@

```{r Lombardo_Comparison}
cvt_DTXSIDs <- unique(my_pk$data$Chemical)
# First things first, let's load the data from Lombardo et al.
lombardo <- readxl::read_xlsx(
  paste0(Sys.getenv("FIG_DIR"),
         "Lombardo2018-Supplemental_82966_revised_corrected.xlsx"),
  skip = 8)

@@

```{r otherMeta-analysis}

@@

```{r do_fit-benchmarking}

@@

```{r sfig6-examples-of-fits}
pl <- plot(my_pk, use_scale_conc = TRUE, n_interp = 10)
names(pl) # Chemical Species observations observation_plot predicted predicted_plot final_plot

@@

```{r more_analysis}

@@

```{r PredictionVariability_overtime}
my_residuals %>%
  filter(exclude == FALSE) %>%
  inner_join(winmodel) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax = tmax.tkstats) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical,
           Species,
           Route,
           Media,
           Dose) %>%
  mutate(
         normTime = ifelse(
           Route == "iv",
           1 + Time/max(Time),
           ifelse(
             Time > tmax,
             ((Time - tmax)/max(Time)) + 1,
             0.5 + Time/(2*tmax)
           )
         )) %>%
  distinct() %>%
  filter(Detect %in% TRUE) -> my_residuals2
my_residuals2 %>%
  # filter(Route == "oral") %>% View()
  ggplot(aes(
    x = normTime,
    y = Residuals
  )) +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID4020533",
                           Species %in% "rat"),
             color = "orange2", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8031865",
                           Species %in% "rat"),
             color = "magenta4", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8025337",
                           Species %in% "rat"),
             color = "green2", size = 0.5) +
  geom_vline(xintercept = 1, color = "blue", linetype = "dashed") +
  geom_vline(xintercept = 2, color = "green4", linetype = "dashed") +
  facet_grid(cols = vars(Route),
             # rows = vars(model),
             scales = "free_x") +
  labs(x = "ADME-Normalized Time",
       y = "Winning Model Residuals") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank())

@@

```{r function_connections}
fw <- mvbutils::foodweb(where = asNamespace("invivoPKfit"),
                        border = TRUE, cex = 0.6, lwd = 1.5)
fun_mat <- fw$funmat

@@

```{r plotting_all_Fits}
### Plotting for all (2.5MB file)
pl <- plot(my_pk, n_interp = 8, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "DoseNormalized_Joint_AllFits_Final.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

@@

### Figure 2
```{r Explain-Heirarchical-Model}
# Note that the data will run one "Series" of time point at a time
my_1comp <- function(t, F_V, kelim, kabs, sd = 0) {
  sd + (F_V * (kabs * (exp(-kelim * t) - exp(-kabs * t))) / (kabs - kelim))
}

@@

```{r cvtdb-prep}

>>>>

```{r pooled_fitting_choices, eval=FALSE}
# Types of Choices:
## Dose Normalized
## Log10 transform
## Scale time

@@

```{r joint_fitting_choices, eval=FALSE}
# Joint
my_pk_000j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

@@

```{r Evaluation-helpers, eval=FALSE}
# Evaluation
# Current:

@@

```{r Evaluating choices, eval=FALSE}

@@

```{r load_multirefs, eval=FALSE}
# Reference Multiples
multiRef_ChemSpec <- cvt_df %>%
  ungroup() %>%
  distinct(analyte_dtxsid, species, document_id) %>%
  group_by(analyte_dtxsid, species) %>%
  count() %>%
  arrange(desc(n)) %>%
  filter(n > 1) %>%
  dplyr::select(!n)
multiRef_ChemSpec %>% nrow() # Surprisingly only 13 Chem + Species have multiple references

@@

```{r Plotting_Summary, eval=FALSE}
finalist_options <- c("000 Pooled", "000 Joint",
                      "100 Pooled", "100 Joint",
                      "101 Pooled", "101 Joint")

@@

```{r Final_Options, eval=FALSE}
my_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species, Reference))

@@

```{r data_aggregation, eval=FALSE}
# Need the following data for any of the subsequent plots

@@

### Figure 4A: Replicated timepoint variation in CvTdb
```{r fig4A, eval=FALSE}

@@

```{r fig4A-2-alt, eval=FALSE}
iad_summary <- cvt %>% filter(conc > 0, n_subjects_in_series == 1) %>%
  group_by(analyte_dtxsid, species, administration_route_normalized,
           conc_medium_normalized, dose_level_corrected, document_id,
           time_hr) %>%
  filter(n() > 1) %>%
  summarize(conc_sd_calc = sd(conc),
            conc_avg_calc = mean(conc),
            n_subjects_in_series = n())

@@

```{r fig4B-cowplot, eval=FALSE}

@@

```{r cvt_analysis, eval=FALSE}
data_range <- all_my_data %>%
  filter(Detect) %>%
  group_by(Chemical, Species, Reference, Media, Route, Dose, Dose.Units) %>%
  summarize(maxT = max(Time)/24,
         concRange = log10(max(Conc)/min(Conc))) %>%
  ungroup()

@@

```{r sFig2-model_performance, eval=FALSE}
my_preds %>%
  right_join(winmodel) %>%
  rowwise() %>%
  mutate(Pred_Obs = Conc_trans/Conc_est) %>%
  filter(is.finite(Pred_Obs),
         model %in% c("model_1comp", "model_2comp")) %>%
  # mutate(pass = ifelse(Pred_Obs < 0.5, TRUE, FALSE)) %>%
  # pull(pass) %>% table()
  ggplot(aes(x = Pred_Obs)) +
  geom_histogram(binwidth = 0.5,
                 fill = "grey5") +
  scale_x_log10(limits = c(NA, 1E6)) +
  expand_limits(y = 0) +
  geom_vline(xintercept = c(0.5, 2), color = "red3", linetype = "dashed") +
  facet_grid(cols = vars(model),
             rows = vars(Route)) +
  labs(x = "Dose-normalized\nPredicted/Observed",
       y = "Number of Observations") +
  theme_classic() +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.4))

@@

```{r fig5-modelPerform-v-dataVar, eval=FALSE}

@@

```{r fig5Alt, eval=FALSE}
# Redo combined summary
combined_summary <- bind_rows(sud_summary, iad_summary) %>%
  rowwise() %>%
  mutate(twofold_95 = ifelse((conc_avg_calc + (2*conc_sd_calc))/conc_avg_calc < 2,
                             TRUE, FALSE))

@@

```{r fig6-goodness-of-fit, eval=FALSE}

@@

```{r sFig3-sigmaDistribution, eval=FALSE}

@@

```{r Lombardo_Comparison, eval=FALSE}
cvt_DTXSIDs <- unique(my_pk$data$Chemical)
# First things first, let's load the data from Lombardo et al.
lombardo <- readxl::read_xlsx(
  paste0(Sys.getenv("FIG_DIR"),
         "Lombardo2018-Supplemental_82966_revised_corrected.xlsx"),
  skip = 8)

@@

```{r otherMeta-analysis, eval=FALSE}

@@

```{r do_fit-benchmarking, eval=FALSE}

@@

```{r sfig6-examples-of-fits, eval=FALSE}
pl <- plot(my_pk, use_scale_conc = TRUE, n_interp = 10)
names(pl) # Chemical Species observations observation_plot predicted predicted_plot final_plot

@@

```{r more_analysis, eval=FALSE}

@@

```{r PredictionVariability_overtime, eval=FALSE}
my_residuals %>%
  filter(exclude == FALSE) %>%
  inner_join(winmodel) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax = tmax.tkstats) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical,
           Species,
           Route,
           Media,
           Dose) %>%
  mutate(
         normTime = ifelse(
           Route == "iv",
           1 + Time/max(Time),
           ifelse(
             Time > tmax,
             ((Time - tmax)/max(Time)) + 1,
             0.5 + Time/(2*tmax)
           )
         )) %>%
  distinct() %>%
  filter(Detect %in% TRUE) -> my_residuals2
my_residuals2 %>%
  # filter(Route == "oral") %>% View()
  ggplot(aes(
    x = normTime,
    y = Residuals
  )) +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID4020533",
                           Species %in% "rat"),
             color = "orange2", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8031865",
                           Species %in% "rat"),
             color = "magenta4", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8025337",
                           Species %in% "rat"),
             color = "green2", size = 0.5) +
  geom_vline(xintercept = 1, color = "blue", linetype = "dashed") +
  geom_vline(xintercept = 2, color = "green4", linetype = "dashed") +
  facet_grid(cols = vars(Route),
             # rows = vars(model),
             scales = "free_x") +
  labs(x = "ADME-Normalized Time",
       y = "Winning Model Residuals") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank())

@@

```{r function_connections, eval=FALSE}
fw <- mvbutils::foodweb(where = asNamespace("invivoPKfit"),
                        border = TRUE, cex = 0.6, lwd = 1.5)
fun_mat <- fw$funmat

@@

```{r plotting_all_Fits, eval=FALSE}
### Plotting for all (2.5MB file)
pl <- plot(my_pk, n_interp = 8, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "DoseNormalized_Joint_AllFits_Final.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

@@

### Figure 2
```{r Explain-Heirarchical-Model, eval=FALSE}
# Note that the data will run one "Series" of time point at a time
my_1comp <- function(t, F_V, kelim, kabs, sd = 0) {
  sd + (F_V * (kabs * (exp(-kelim * t) - exp(-kabs * t))) / (kabs - kelim))
}

@@

```{r cvtdb-prep, eval=FALSE}
